---
title: "Rainfall analysis"
author: "Benjamin Chan"
date: `r Sys.Date()`
output:
  html_document:
    keep_md: yes
---


Forked from Jupyter Notebook https://github.com/engjen/rain_pdx

Load packages

```{r}
library(magrittr)
library(tidyverse)
library(ggthemes)
```

Read rain gauge data from City of Portland Bureau of Environmental Services by way of USGS.

> Vernon School Rain Gage - 2044 NE. Killingsworth St.  
>  
> PROVISIONAL, UNCORRECTED RAW DATA FROM THE CITY OF PORTLAND HYDRA NETWORK.  
> Data are the number of tips of the rain gage bucket.  
> Each tip is 0.01 inches of rainfall.  
>  [`-`, missing data]  
> Dates and times are PACIFIC STANDARD TIME.  

Assume most recent day is incomplete and exclude.
Convert *tips* to inches.

```{r}
url <- "https://or.water.usgs.gov/non-usgs/bes/vernon.rain"
f <- tempfile()
download.file(url, f)
colNames <- read_lines(f, skip = 9, n_max = 1) %>% str_trim() %>% str_split("\\s+") %>% unlist()
L <- read_lines(f, skip = 11) %>% str_split("\\s+")
date <- sapply(L, function(x) x[1]) %>% as.Date(format = "%d-%b-%Y")
total <- sapply(L, function(x) x[2]) %>% as.numeric()
hourly <- sapply(L, function(x) x[3:26], simplify = FALSE) %>% do.call(rbind, .) %>% data.frame() %>% mutate_if(is.character, as.numeric)
names(hourly) <- paste0("x", colNames[3:26])
df <-
  tibble(date, hourly) %>%
  filter(date < max(date)) %>%
  pivot_longer(starts_with("x")) %>%
  mutate(hour = gsub("x", "", name) %>% as.integer()) %>%
  select(-name) %>%
  mutate(datetime = sprintf("%02d:00", hour) %>% paste(date, .) %>% as.POSIXlt(format = "%Y-%m-%d %H:%M")) %>%
  mutate(inches = value * 0.01) %>%
  select(-value)
head(df) %>% knitr::kable()
```

Plot the most recent 5 years

```{r}
y <- 5
G <-
  df %>%
  filter(max(date) - y * 365.25 <= date) %>%
  mutate(year = year(date) %>% factor(),
         month = month(date),
         week = week(date)) %>%
  group_by(year, week) %>%
  summarize(inches = sum(inches)) %>%
  ungroup() %>%
  ggplot(aes(x = week, y = inches, color = year)) +
    geom_line(size = 1, show.legend = FALSE) +
    scale_x_continuous("Week") +
    scale_y_continuous("Inches of rainfall", transform = "sqrt") +
    scale_color_brewer(type = "qual") +
    # coord_polar() +
    facet_wrap(~ year, ncol = 1)
G
```

Create `ts` object

```{r}
min <- min(df$date, na.rm = TRUE)
max <- max(df$date, na.rm = TRUE)
timeSeries <-
  df %>%
  group_by(date) %>%
  summarize(inches = sum(inches)) %>%
  ungroup() %>%
  ts(start = min, end = max)
```
